\documentclass{article}

\usepackage[utf8]{inputenc}

\usepackage[obeyDraft]{todonotes}

\pagestyle{empty}

\begin{document}
\begin{center}
\Large{\sc Application for DL 2017 grant}
\end{center}

\bigskip
\noindent Dear DL 2017 chairs,

\bigskip

\noindent
hereby I would like to apply for a student grant for participation in DL 2017, that will be organized in Montpellier, France, July 18-21, 2017, since it will help to cover my costs related to participation in DL.

\bigskip
\noindent Bratislava, 16 May 2017,\\
J\'{u}lia Pukancov\'{a}

\newpage

\begin{center}
\Large{\sc Personal Details}
\end{center}

\bigskip

\begin{tabular*}{\textwidth}{ l l }
 \textbf{Name} & Júlia Pukancová \\ [1em]
 \textbf{Email} & pukancova@fmph.uniba.sk \\ [1em]
 \textbf{Institution} & Comenius University in Bratislava, \\
 & Faculty of Mathematics, Physics and Informatics \\ 
 & Mlynská dolina F1 \\
 & 842 48 Bratislava\\
 & Slovakia \\[1em]
 \textbf{Current level of studies} & doctoral (PhD) \\[1em]
 \textbf{Thesis topic} & Abductive Diagnosis with Description Logics \\[1em]
 \textbf{Supervisor} & Ján Šefránek \\[1em]
  \textbf{DL 2017 submission} & yes (paper)
\end{tabular*}

\newpage

\begin{center}
\Large{\sc Statement of Motivation}
\end{center}

\bigskip

\noindent
I would like to attend the 30th Workshop  on Description Logics, held in Montpellier, France, 18-21 July, 2017 because it is highly relevant for my PhD research.

My dissertation project is focused on abductive reasoning in description logics. I have attended DL~2016 and DL~2015, both with submitted papers. At DL~2016 I have presented our ABox abduction algorithm for DL. Recently, we have extended this algorithm and I would like to present our current results at DL~2017. I have very good experience with DL workshop, as it has always enriched my research. Moreover, at both DL workshops I got interesting feedback which I appreciate a lot. Unfortunately, my faculty is not able to completely cover my travel costs.

\bigskip
\noindent
Júlia Pukancová

\newpage

\begin{center}
\Large{\sc Letter from Supervisor}
\end{center}

\bigskip
\noindent To whom it may concern,

\bigskip

\noindent I would like to express my sincere support to my master student Katarína Fabianová in respect to her application for the students grant for attending
DL~2019.

Katarína is a full time second-year master student of our master program in
applied informatics. In her diploma thesis she has researched approaches suitable for abductive reasoning in DL. Therefore her work is relevant for the DL workshop. It would be very helpful for her to present the recent results and to get some feedback from the DL community.

Unfortunately the funding available for master students at our university is
rather limited, and therefore without the DL students grant she will not be able to attend the workshop.

\bigskip
\noindent Kind regards,

\bigskip
\noindent Júlia Pukancová\\
Dept. of Applied Informatics\\
Faculty of Mathematics, Physics, and Informatics\\
Comenius University in Bratislava

\end{document}
